---
title: Canary Current LME robis demo
date: "`r Sys.Date()`"
author: Pieter Provoost
output: (function(...) {
  rmdformats::robobook(toc_depth = 4, pandoc_args = c("+RTS", "-K2000m", "-RTS"), ...) })
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })  
---

This notebook is a quick demonstration of the robis R package for fetching occurrence data relevant to the Canary Current LME.

## Dependencies

```{r message=FALSE, warning=FALSE}
library(robis)
library(dplyr)
library(ggplot2)
library(sf)
library(rnaturalearth)
```

## Fetching occurrences

The OBIS database has a number of geographic areas against which the occurrence data are indexed. This includes LMEs. Use the `areas()` function to get a list of LME area identifiers.

```{r message=FALSE, warning=FALSE}
areas <- area() %>%
  filter(type == "lme")

areas %>% rmarkdown::paged_table()
```
<br/>
Then use the Canary Current LME area ID as a filter in the `occurrence()` function.

```{r message=FALSE, warning=FALSE, cache=TRUE}
occ <- occurrence(areaid = 40027)
occ
```

Let's create a map:

```{r message=FALSE, warning=FALSE, cache=TRUE}
land <- ne_coastline(returnclass = "sf", scale = "large")

ggplot() +
  geom_sf(data = land, size = 0.5) +
  geom_sf(data = occ %>% st_as_sf(coords = c("decimalLongitude", "decimalLatitude"), crs = 4326), size = 0.1, color = "#45add6") +
  coord_sf(xlim = c(-26, -3), ylim = c(10, 38), default_crs = sf::st_crs(4326)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

## WRiMS taxa

While it's possible to filter the dataframe we downloaded earlier based on value in the `wrims` column, the `occurrence()` function also has a `wrims` filter which restricts results to species listed in the World Register of Introduced Marine Species (WRiMS). Note that this currently does not take into account native versus introduced ranges.

```{r message=FALSE, warning=FALSE, cache=TRUE}
occ_wrims <- occurrence(areaid = 40027, wrims = TRUE)
occ_wrims
```

```{r message=FALSE, warning=FALSE, cache=TRUE}
ggplot() +
  geom_sf(data = land, size = 0.5) +
  geom_sf(data = occ_wrims %>% st_as_sf(coords = c("decimalLongitude", "decimalLatitude"), crs = 4326), size = 0.1, color = "#45add6") +
  coord_sf(xlim = c(-26, -3), ylim = c(10, 38), default_crs = sf::st_crs(4326)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Let's take a look at the most common WRiMS species for the CCLME:

```{r message=FALSE, warning=FALSE, cache=TRUE}
occ_wrims %>%
  filter(!is.na(species)) %>%
  group_by(class, species) %>%
  summarize(records = n()) %>%
  arrange(desc(records)) %>%
  rmarkdown::paged_table()
```

